# AR models 


Let's reproduce the AR simulation in presentation 

## AR(1) simulation

Let's simulate AR(1)

```{r}
set.seed(1)
y.ar <- e <- rnorm(100)
for (t in 2:100) y.ar[t]  <- 0.7 * y.ar[t-1] +e[t]
```


```{r, out.width="70%"}
plot(y.ar,type="l")            
```



```{r, out.width="100%"}
acf(y.ar)
```
           

```{r , out.width="100%"}
pacf(y.ar)
```


## Model fitted to simulated AR(1)

Let's use the `ar()` function to our simulated data 

```{r}
y.ar.mod <- ar(y.ar, method="mle")
y.ar.mod$order
```
The model correctly find the right order 1

```{r}
y.ar.mod$ar
```

Let's find the confidence interval for the coefficient

```{r}
y.ar.mod$ar + c(-2,2) * sqrt(y.ar.mod$asy.var)
```

```{r}
y.ar.ts <- ts(y.ar, start=1980, frequency = 12)
auto.arima(y.ar.ts)
```

## Reproduce with simts package


```{r}
set.seed(1)
true_model=AR(phi=0.70,sigma2=1) 
y.ar.sim <- gen_gts(n=100,model=true_model)
```

Let's have a look at pacf

```{r}
acf(y.ar.sim)
pacf(y.ar.sim)
```

Then use the select method 

```{r}
select(AR(10), y.ar.sim, include.mean=F)
```

We find that all the criteria selected lag of order 1

Let's estimate that model

```{r}
y.est <- estimate(AR(1), y.ar.sim)
```

Now build the confidence interval 

```{r}
y.est$mod$coef[["ar1"]] + c(1.96,-1.96) * sqrt(y.est$mod$var.coef[1,1])
```

We find that the confidence interval contains the true value



Now let's try another model which AR(3)

```{r}
set.seed(1)
true_model = AR(phi = c(1.2, -0.9, 0.3), sigma2 = 1)
y.ar3 = gen_gts(n = 500, model = true_model)
```

Let's have a look at acf and pacf

```{r}
acf(y.ar3)
pacf(y.ar3)
```

Now let's use select 

```{r}
select(AR(10), y.ar3, include.mean = F)
```

In the above we find that all of the measures are selecting the 3

```{r}
y.ar3.est <- estimate(AR(3), y.ar3)
y.ar3.est$mod$coef
var.coef <- diag(y.ar3.est$mod$var.coef)[1:3]
sd.coef.mat <- matrix(sqrt(var.coef),ncol=1)
```

Let's find some matrices 


```{r}
crit <- matrix(c(-2,2),nrow=1)
coefs <- rep(y.ar3.est$mod$coef[1:3],2) %>% 
                matrix(ncol=2)
coefs + sd.coef.mat %*% crit
```

Let's apply auto.arima

```{r}
y.ar3.ts <- ts(y.ar3, start=1980, frequency = 12)
auto.arima(y.ar3.ts)
```

Above auto.arima() function is no longer needed. It has been replaced by `fable` package `ARIMA` function


Let's try `tidyverts` framework

```{r}
y.ar3.tsl <- as_tsibble(y.ar3.ts)
fit  <- y.ar3.tsl %>% 
          model(ARIMA(value ~  PDQ(0,0,0)))
report(fit)

y.ar3.tsl <- as_tsibble(y.ar3.ts)
fit2  <- y.ar3.tsl %>% 
          model(ARIMA(value ~ pdq(3,0,0)+ PDQ(0,0,0)))
report(fit2)
```

According to documentation, pdq() specifies non-seasonal order whereas
PDQ() 


There seems to be arima.sim() function in `stats` packcage

```{r}
set.seed(1)
rw <- arima.sim(list(order = c(0,1,0)), n = 200)
wn <- arima.sim(list(order = c(0,0,0)), n = 200)
ar1 <- arima.sim(list(order = c(1,0,0),ar=0.7), n = 200)

ar1.tsb <- as_tsibble(ar1)

ar1.est <- ar1.tsb %>% 
                model(AR(value ~ order(1)))
ar1.est  %>% 
    coef()

ar1.est  %>% 
    tidy()

ar1.est  %>% 
    glance()


ar1.est  %>% 
    hilo()
```


## Time series observations simulation 


### AR1 whole distribution simulation

```{r}
ar1.s100 <- do(100) * arima.sim(list(order = c(1,0,0),ar=0.7), n = 1000)
rownames(ar1.s100)=paste0("sim",rownames(ar1.s100))
names(ar1.s100)=paste0("t",1:ncol(ar1.s100))

ar1.s100.t <- ar1.s100 %>%
   tibble::rownames_to_column() %>%  
   pivot_longer(-rowname) %>% 
   pivot_wider(names_from=rowname, values_from=value) 

ar1.s100.t %>% 
    rowwise() %>% 
    mutate(mu=mean(c_across(sim1:sim100))) %>% 
    select(time=name, mu)  -> ar1.s100.mu

ar1.s100.mu %>% 
    ggplot() +
      geom_line(aes(x=time,y=mu), group="") +
      ylim(-1,1)
```

What we find is that mean is not constant

Now we rather try to generate a normal distribution with known mean

### Normal distribution whole distribution simulation

```{r}
norm.s100 <- do(100) * rnorm(1000, mean=5, sd=1) 
rownames(norm.s100)=paste0("t",1:nrow(norm.s100))
names(norm.s100)=paste0("sim",1:ncol(norm.s100))
norm.s100.t <- norm.s100 %>%
   tibble::rownames_to_column() %>%  
   pivot_longer(-rowname) %>% 
   pivot_wider(names_from=rowname, values_from=value) 

norm.s100 %>% 
    rowwise() %>% 
    transmute(mu=mean(c_across(sim1:sim100))) %>% 
    rownames_to_column("time") -> norm.s100.mu

norm.s100.mu %>% 
    ggplot() +
      geom_line(aes(x=time,y=mu), group="") +
      ylim(0,10)
```

What we find is that mean is not constant

Now we rather try to generate a normal distribution with known mean




# VAR from ITSR 

```{r}
library(tidyverse)
library(tsibble)
library(feasts)
```


Chpater 11 in ITSR

```{r}
www <- "~/github_repos/Times-series/Misc/itsr_data/cbe.dat"
cbe <- read.table(www, header = T)
```

Let's convert this into tsibble 

```{r}
cbe.ts <- ts(cbe, start=1958, freq=12) %>% 
             as_tsibble()
```

Let's plot 

```{r}
cbe.ts %>% 
    autoplot(value)
```

Here is 

```{r}
library(mvtnorm)
```
Here we go

```{r}
cov.mat <- matrix(c(1, 0.8, 0.8, 1), nr = 2)
w <- rmvnorm(1000, sigma = cov.mat)
cov(w)
wx <- w[,1]
wy <- w[,2]
```

We create wn series with simts 

```{r}
x.wn <- gen_gts(1000,WN(sigma2=1))
y.wn <- gen_gts(1000,WN(sigma2=1))
wn <- gen_gts(1000,WN(sigma2=1))
var.x=0.4*lag(x.wn)+0.3*lag(y.wn)+ wx
var.y=0.2*lag(x.wn)+0.1*lag(y.wn)+ wy
var.xy <- cbind(var.x, var.y)
colnames(var.xy)  <- c("x","y")
var.xy <- var.xy[2:nrow(var.xy),]
xy.ar <- ar(var.xy)
xy.ar$ar
```
Let's try this 

```{r}
x.ar <- gen_gts(1000,AR1(phi=.32,sigma2=1))
y.ma <- gen_gts(1000,MA1(theta=.43, sigma2=1))
wn <- gen_gts(1000,WN(sigma2=1))
model.x <- .4*AR1(phi=.32,sigma2=1)+.5*MA1(theta=.43, sigma2=1)+ WN(sigma2=1)
var.x <- gen_lts(1000,model=model.x)
var.y=0.2*lag(x.ar)+0.1*lag(y.ma)+ wn
var.xy <- cbind(var.x, var.y)
colnames(var.xy)  <- c("x","y")
var.xy <- var.xy[2:nrow(var.xy),]
xy.ar <- ar(var.xy)
xy.ar$ar
```

Not very successful, let's try this  using library `tsDyn` package




```{r}
library(tsDyn)
```

Now as we will use the `VAR.sim()` function from that package
It implements the VAR model in Enders

It starts from the structural VAR but then moved onto standard VAR 

$$y_t=a_{10} + a_{11}y_{t-1}+a_{12}z_{t-1}+e_1t$$
$$z_t=a_{20} + a_{21}y_{t-1}+a_{22}z_{t-1}+e_2t$$


```{r}
a10 <- a20 <- 0
a11 <- a22 <- 0.7
a12 <- a21 <- 0.2
B1<-matrix(c(a11, a12, a21, a22), 2)
var1<-VAR.sim(B=B1,n=100,include="none")
colnames(var1) <- c("yt","zt")
ts.plot(var1, type="l", col=c(1,2))
```


Let's estimate

```{r}
var.aic <- VAR(var1, type = "none", lag.max = 5, ic = "AIC")
```

Let's compare the original parameter and estimated values

Original parameters

```{r}
B1
```

Estimated parameters

```{r}
coef(var.aic)
```

By comparison we find that the estimates are pretty close.

Let's check other functions from `vars` package

```{r}
VARselect(var1)
```

The above selects 1 lag length according to SC which is more parsimonious. So we will follow that.


Now let's try to reproduce the following from 

`https://www.r-econometrics.com/timeseries/varintro/`


```{r}
set.seed(123) # Reset random number generator for reasons of reproducability
# Generate sample
t <- 200 # Number of time series observations
k <- 2 # Number of endogenous variables
p <- 2 # Number of lags

# Generate coefficient matrices
A.1 <- matrix(c(-.3, .6, -.4, .5), k) # Coefficient matrix of lag 1
A.2 <- matrix(c(-.1, -.2, .1, .05), k) # Coefficient matrix of lag 2
matA <- cbind(A.1, A.2) # Companion form of the coefficient matrices
var2<-VAR.sim(B=matA,lag=2,n=200,include="none")
colnames(var2) <- c("yt","zt")
ts.plot(var2, type="l", col=c(1,2))
```
Now let's check whether it selects the right number of lags

```{r}
VARselect(var2)
```

Let's now estimate


```{r}
var.aic <- VAR(var2, type = "none", lag.max = 5, ic = "AIC")
```

Time for `IRF`


```{r}
ir.1 <- irf(var.aic, impulse = "yt", response = "zt", n.ahead = 20, ortho = FALSE)

# Plot the IRF
plot(ir.1)
```




Now let's do ITSR

```{r}
x <- y <- rep(0, 1000)
x[1] <- wx[1]
y[1] <- wy[1]
for (i in 2:1000) {
    x[i] <- 0.4 * x[i - 1] + 0.3 * y[i - 1] + wx[i]
    y[i] <- 0.2 * x[i - 1] + 0.1 * y[i - 1] + wy[i]
}
xy.ar <- ar(cbind(x, y))
xy.ar$ar[, , ]
```



