class: inverse,middle,center

# Mixed models: The ARMA process 

---

## Definition

An AR(p) process can be described in the following manner: 

$$x_t=\alpha_1x_{t-1}+\alpha_2x_{t-2}+\ldots+\alpha_px_{t-p}+w_t$$ 

Here ${w_t}$ is white noise and the $\alpha_i$ are the model parameters. 

A useful class of models are obtained when AR and MA terms are added together in a single expression 

A time series ${x_t}$ follows an autoregressive moving average (ARMA) process of order $(p,q)$, denoted $ARMA(p,q)$ when

$$\begin{equation}
x_t=\alpha_1 x_{t-1} + \alpha_2 x_{t-2}+\ldots +\alpha_p x_{t-p}+w_t +\beta_1 w_{t-1} + \beta_2 w_{t-2}+\ldots+\beta_q w_{t-q}
\end{equation}$$


Here ${w_t}$ is white noise 

---

The above equation can be represented in terms of the backward shift operator `B` and rearranged in the more concise polynomial form 

$$\theta_p(B)x_t=\phi_q(B)w_t$$ 

The following points should be noted about an ARMA(p,q) process: 

The process is stationary when the roots of $\theta$ all exceed unity in absolute value 

The process is invertible when the roots of $\phi$ all exceed unity in absolute value 

The AR(p) model is the special case $ARMA(p,0)$ 

The MA(q) model is the special case ARMA(0,q) 

Parameter parsimony :  When fitting to data, an ARMA model will often be more parameter efficient - require fewer parameters - than a single MA or AR model 

Parameter redundancy: when $\theta$ and $\phi$ share a common factor, a stationary model can be simplified. For example, the model $(1-\frac{1}{2}B)(1-\frac{1}{3})x_t=(1-\frac{1}{2})w_t$ can be written as $(1-\frac{1}{3})x_t=wt$ 


---

## ARMA models: Empirical analysis

**Simulation and fitting**

The ARMA process can be simulated using the R function `arima.sim`

This function takes a list of coefficients representing the AR and MA parameters. 

An ARMA(p,q) model can be fitted using the `arima` function with the `order` function parameter set to `c(p,0,q)` 

The fitting algorithm proceeds similarly to that for a MA process. 

The fitting algorithm proceeds similarly to that for an MA process 

Below data from an ARMA(1,1) process are simulated for $\alpha$=-0.6  and $\beta$=0.5

As expected, the sample estimates of $\alpha$ and $\beta$ are close to the underlying model parameters 

```{r }
set.seed(1)
x <- arima.sim(n = 10000, list(ar = -0.6, ma = 0.5))
coef(arima(x, order = c(1, 0, 1)))
```
---

## Exchange rate series

Previously, we have found that a simple MA(1) model failed to provide an adequate fit to the exchange rate series. 

In the code below, fitted MA(1), AR(1) and ARMA(1,1) models are compared using the AIC 

ARMA(1,1) model provides the best fit to the data, followed by AR(1) model, along with MA(1) model providing the poorest fit. 

The figure indicates that the residuals of the fitted ARMA(1,1) model have small autocorrelations

This is consistent with a realisation of white noise and supports the use of the model 

---

```{r, eval=FALSE }
x.ma <- arima(x.ts, order = c(0, 0, 1))
x.ar <- arima(x.ts, order = c(1, 0, 0))
x.arma <- arima(x.ts, order = c(1, 0, 1))
```

Now let's find the AIC's 

```{r, eval=FALSE }
AIC(x.ma)
AIC(x.ar)
AIC(x.arma) 
acf(resid(x.arma))
```

---

## Electricity production series

Let's consider the Australian electricity production series 

The data exhibit a clear positive trend and a regular seasonal cycle 

The variance increases with time

This suggests a log-transformation may be appropriate 

---

A regression model is fitted to the logarithms of the original series in the code below 

```{r, eval=FALSE }
www <- "http://www.massey.ac.nz/~pscowper/ts/cbe.dat"
CBE <- read.table(www, header = T)
Elec.ts <- ts(CBE[, 3], start = 1958, freq = 12)
Time <- 1:length(Elec.ts)
Imth <- cycle(Elec.ts)
Elec.lm <- lm(log(Elec.ts) ~ Time + I(Time^2) + factor(Imth))
acf(resid(Elec.lm))
```

---

The correlogram of the residuals appears to cycle with a period of 12 months suggesting that the monthly indicator variables are not sufficient to account for the seasonality in the series 

When we study ARIMA, we will find that this can be accounted for using a non-stationary model with a stochastic seasonal component. 

In the meantime, the best fitting $ARMA(p,q)$ model can be chosen using the smallest AIC either by trying a range of combinations of $p$ and $q$ in the `arima` function

Or using a `for` loop with upper bounds on $p$ and $q$ 

In each step of the `for` loop, the AIC of the fitted model is commpared with the currently stored smallest value 

---

If the model is found to be an improvement, that is it has smaller AIC value, then the new value and model are stored 

To start with, best.aic is initialised to infinity $Inf$ 

After the loop is complete, the best model can be found in best.order 

In this case, the best model turns out to be an AR(2) model 

```{r, eval=FALSE }
best.order <- c(0, 0, 0)
 best.aic <- Inf
 for (i in 0:2) for (j in 0:2) {
fit.aic <- AIC(arima(resid(Elec.lm), order = c(i, 0,
j)))
if (fit.aic < best.aic) {
best.order <- c(i, 0, j)
best.arma <- arima(resid(Elec.lm), order = best.order)
best.aic <- fit.aic
}}
```

---

Now let's see the best order 

```{r, eval=FALSE }
best.order 
```

Now time for the ACF

```{r, eval=FALSE }
acf(resid(best.arma))
```

---

The `predict` function can be used to both to forecast future values from the fitted regression model

Forecast the future errors associated with the regression model suing the ARMA model fitted to the residuals from the regression 

These two forecasts can then be summed to give a forecasted value of the logarithm for electricity production

This would need to be anti-logged and perhaps adjusted using a bias correction factor 

`predict` is a generic `R` function, it works in different ways for different input objects and classes 

For a fitted regression model of class `lm`, the `predict` function requires the new set of data to be in the form of a data frame 

For a fitted ARMA model of class `arima`, the `predict` function requires just the number of time steps ahead for the desired forecast 

In the latter case, `predict` produces an object that has both the predicted values and their standard errors

These two values can be extracted using `pred` and `se`

---

In the code below, the electricity production for each month of the next three years is predicted 

```{r, eval=FALSE }
 new.time <- seq(length(Elec.ts), length = 36)
 new.data <- data.frame(Time = new.time, Imth = rep(1:12,
3))
 predict.lm <- predict(Elec.lm, new.data)
 predict.arma <- predict(best.arma, n.ahead = 36)
 elec.pred <- ts(exp(predict.lm + predict.arma$pred), start = 1991,
freq = 12)
 ts.plot(cbind(Elec.ts, elec.pred), lty = 1:2)
```

The plot of the forecasted values suggests that the predicted values for winter may be underestimated by the fitted model 

This may be due to the remaining seasonal autocorrelation in the residuals 

This problem will be tackled in ARIMA topic 

---

## Wave tank data

The data in the file `wave.dat` are the surface height of water (mm), relative to the still water level 

This is measured using a capcitance probe positioned at the centre of a wave tank 

The continuous voltage signal from this capacitance probe was sampled every 0.1 second over a 39.6 second period. 

The objective is to fit a suitable ARMA(p,q) model that can be used to generate a realsistic wave input to a mathematical model for an ocean-going tugboat in a computer simulation 

The results of the computer simulation will be compared with tests using a physical model of the tugboat in the wave tank 

The pacf suggest that $p$ should be at least 2 

---

The best fitting ARMA(p,q) model based on a minimum variance of residuals was obtained with both $p$ and $q$ equal to 4. The acf and pacf of the residuals from this model are consistent with the residuals being a realisation of white noise. 

```{r, eval=FALSE }
www <- "http://www.massey.ac.nz/~pscowper/ts/wave.dat"
wave.dat <- read.table(www, header = T)
attach (wave.dat)
layout(1:3)
plot (as.ts(waveht), ylab = 'Wave height')
acf (waveht)
pacf (waveht)
wave.arma <- arima(waveht, order = c(4,0,4))
acf (wave.arma$res[-(1:4)])
pacf (wave.arma$res[-(1:4)])
hist(wave.arma$res[-(1:4)], xlab='height / mm', main='')
```









